Negative Perceptions of Outsourcing to Artificial Intelligence

### 2025-07-15 manuscript.qmd
Make further manuscript edits

---
title: "Negative Perceptions of Outsourcing to Artificial Intelligence"
shorttitle: "Outsourcing to AI"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    email: pv201@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - formal analysis
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: This pre-print is currently not yet peer-reviewed, and may differ from the final version.
    data-sharing: Word count = 10,574 words.
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: As artificial intelligence (AI) tools become increasingly integrated into daily life, people are beginning to outsource not only professional tasks but also socio-relational ones. Large language models like ChatGPT can generate wedding vows, speeches, and personal messages, raising questions about how individuals who use AI for such tasks are perceived by others. In this paper, we conduct six pre-registered studies with British participants (N = 3,935) to understand how people view those who expend less effort by outsourcing tasks to AI in different ways. We show that outsourcing makes us think more negatively about not only the person and their motivations, but also the outsourced work itself. Advancing previous work on effort, we show that the specific kind of task that people outsource -- and therefore expend less effort on -- matters; that reduced effort does not lead to uniformly negative character perceptions across different tasks; and that effort is important because it signals second-order judgments that people are being less authentic and care less about the task. Our research highlights the way that AI use shapes our perceptions of people, raising key philosophical questions about efficiency, authenticity, and social ties in a world filled with AI-mediated interactions.
keywords: [artificial intelligence, person perception, outsourcing, effort, trust]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-docx:
    fig-dpi: 300
    fig-align: center
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
---

@@

The widespread release of generative AI language models has transformed daily 
life, offering the potential to perform a variety of tasks more efficiently and,
in some cases, with greater effectiveness than by doing them oneself. But as AI
becomes more widely available, people are not only using it to assist them with
things like preparing dinner recipes, writing data analysis code, and planning
daily schedules. Increasingly, AI might be used beyond routine or technical
domains to instead assist in tasks that are more socio-relational in nature,
like writing wedding vows, apology notes, and love letters. Anecdotal evidence
suggests that not only is AI-outsourcing of this kind already happening, but
that it potentially has serious effects on how we judge others. In a recent
Reddit post, a disgruntled newlywed tells the story of her husband using ChatGPT
to write his wedding vows, expressing her discomfort with outsourcing something
to AI that, to her, is deeply meaningful and a reflection of their love for one 
another [@miramar0]. Other reports describe people's negative reactions when
they learn that their romantic partner used ChatGPT to write them an apology 
note [@Tait2024] or even to break up with them [@Anderson2025]. Outsourcing 
tasks -- especially socio-relational ones -- to AI tools may be efficient, but 
could have negative consequences for person perception.

There is nothing new, in principle, about outsourcing tasks. For hundreds of 
years, personal assistants have organized daily schedules, recipe-books have 
provided meal plans, and guidebooks have created travel itineraries. In the 
socio-relational domain, ghostwriters have long-existed, and the internet is 
abound with professional paid services for writing wedding vows and personal 
speeches. AI merely supercharges what is an ancient human impulse: the push to 
reduce mental energy by outsourcing parts of our work onto people, books, tools,
or systems. But even if outsourcing is an old phenomenon, the rapid shift in 
availability and use of AI models has fundamentally changed the ease with which
people can outsource work, what kinds of tasks they can outsource, and the way
in which they can outsource. These new developments in society mean that even as
an old phenomenon in new clothes, there is much we still need to know about
outsourcing.

First, we need to know how people who outsource tasks to AI specifically are 
perceived. We know that people are increasingly using large language models 
(LLMs) for a wide variety of tasks [@UKGov2024]. Due to their ubiquity, perhaps 
outsourcing to LLMs might not lead to negative perceptions? We are sceptical. We 
know that people dislike it when others "free ride" or reduce effort while 
benefiting from collective resources [e.g. @Cubitt2011; @Kerr1983] and that 
people's outputs are perceived as more valuable the more effort was ostensibly 
put into them [@Kruger2004]. Moreover, exertion of effort is deemed morally 
admirable and is rewarded, even in situations where effort does not directly 
generate additional product, quality, or economic value, suggesting that effort 
itself is moralized [@Celniker2023]. While outsourcing and reduced effort are
not synonymous, it seems the core social psychological processes are likely to
be similar: when someone outsources a task instead of doing it themselves, they
are expending less effort to achieve their goal, and people value effort. 
Indeed, some work shows that describing someone as using AI for a relational 
task led to the perception they expended less effort and were less satisfied 
with their relationship [@Liu2024]. Other work finds that people receive more 
negative social evaluations if they use AI to complete assignments at school and
at work [@Reif2025; @Roth2025]. We do not know, however, how outsourcing -- and 
therefore, presumably, expending less effort -- might be different for a human 
or an AI, and how these processes will occur in the myriad ways in which people 
can and do use AI to help them complete tasks.

Second, we need to know whether the *type* of task that people are outsourcing 
matters. One might expect outsourcing to be perceived negatively regardless of 
the type of task being outsourced -- if effort is generally moralized, then the 
domain in which it is expended (or not) should have little impact. However, 
there are reasons to expect differences between social tasks like writing vows 
and non-social tasks like writing computer code. We know that different norms,
standards, and expectations can be applied to social and non-social tasks and 
exchanges [e.g. @Fiske1992; @Heider1958; @Malle2022]. Moreover, from a 
philosophical perspective, it often matters not only *whether* something is 
done, but *how* it is done [@Aristotle2009; @Hursthouse2023; @Stohr2006]. An 
apology is not just about hearing someone say "I am sorry", but seeing genuine 
regret; a love letter is not just about hearing someone say "I love you", but 
seeing depth of emotion; and a bereavement letter is not just about hearing 
someone say "I am sorry for your loss", but seeing an understanding for the 
powerful human experience of loss. There is, perhaps especially for social 
tasks, value not only in the outcome of doing something, but the *process* too 
[@Goodman2010]. Previous work on the role of effort, however, has tended to 
focus on a limited number of tasks, looking at how people judge the quality of 
products like paintings, poems, or suits of armour based on how much effort is 
perceived [@Kruger2004] or how people judge others who expend more or less 
effort in employment and physical effort like marathon running [@Celniker2023]. 
Yet the number of tasks that people can outsource –- and therefore, presumably, 
expend less effort –- is vast, and these tasks can vary in different ways. 
Different kinds of tasks that people can outsource can be more or less socially 
relevant, more or less consequential, and more or less likely to impact others: 
it is likely that whether someone puts effort into a socially-relevant task like 
planning a friend's birthday party or writing an apology letter would have 
stronger effects on perceived morality than whether they put effort into writing 
computer code or planning a daily schedule – yet this reminds untested. To 
understand any potential negative effects of outsourcing -- to AI or to humans 
-- we must therefore look at a broad range of non-social and social tasks, 
rather than draw broad conclusions based on a few use cases.

Third, we need to know how outsourcing may shape different kinds of social 
perceptions. People can judge others on separate dimensions of warmth and 
competence [e.g., @Abele2021; @Fiske2007] as well as on dimensions of morality 
and trustworthiness [@Goodwin2014]. It remains unclear how outsourcing to AI 
might lead to differential character judgments across these different 
dimensions. For example, @Kruger2004 focus on judgments of the output, but not 
the person, and while @Celniker2023 include the core social dimensions of 
warmth, morality, and trustworthiness, we do not know how these character 
ratings would be affected by outsourcing to an AI (as well as a human) and how 
this might differ from laziness as a distinct trait. And, most importantly, we 
do not know how outsourcing different kinds of tasks might influence different 
perceptions: does outsourcing in the social domain lead to particularly negative 
effects on warmth, morality, and trustworthiness?

Fourth, we need to know how different *ways* of outsourcing to AI influence 
negative perceptions. Someone who "fully" outsources a task to AI by simply 
giving it a prompt and copying the output word-for-word might be perceived very
differently to someone who gives the AI a prompt, revises the work accordingly,
and finishes it themselves -- using AI as a *collaborative tool*, rather than as
a replacement. Similarly, someone could deceive others about their use of AI or
be perfectly honest about it, and someone could use AI because they are in a 
rush or because the task is important to them and they want to get it right. 
While it seems reasonable to assume that "fully" outsourcing would be perceived 
worse than using AI as a collaborative tool, and that not acknowledging AI use 
would be perceived worse than being honest about it, it remains unclear how much 
this reduces negative perceptions: if someone uses AI in the "best" way (by 
using it as a collaborative tool, being open about their use of AI, and having a 
good reason for using AI), would they still suffer negative social consequences 
from doing so?

Fifth, we need to understand *why* outsourcing to AI, and therefore expending 
less effort, might have these effects. Previous work has focused on how 
expending less effort directly leads to negative perceptions of others 
[@Celniker2023]. But this raises the question of *why* effort is seen as 
important and what exactly it is signalling to others, beyond one's general 
cooperative intent. It is possible that outsourcing leads to negative 
perceptions because the lack of effort spent on the task signals something more 
fundamental about how authentic one is and how much one cares about the task: 
when someone chooses to outsource a love letter to an AI, they might be seen as 
valuing that love letter and what it represents less. It could be this 
second-step order of perceptions that is the key driver of negative perceptions, 
and this, in turn, could especially be so for the kinds of tasks that are more 
socially relevant.

@@

In this paper, we build on classic social psychological work on character 
inferences from reduced effort to understand how people view others who
outsource different kinds of tasks, in different ways, for different reasons, to
AI. Across six pre-registered experiments with British participants, we seek
not only to understand how reduced effort through AI-outsourcing might shape
perceptions of others, but also advance previous theoretical work on the
moralization of effort [@Celniker2023]. Moving beyond a focus on how effort 
itself shapes character judgments, we seek to provide deeper theoretical nuance 
into why effort is important: not only because effort itself is intrinsically 
moralized, but that expending effort gives rise to second-order inferences that 
independently influence character judgments.

@@

Interestingly, we found no effect of authenticity as proxied by the use of a 
personalized AI that is trained on one's own prior writings compared to a 
standard AI like ChatGPT. There are different explanations of this: it is 
possible that authenticity is not an important mechanism underlying the effect 
of outsourcing on negative character evaluations; it is possible that 
authenticity is important but our specific manipulation did not move the needle 
on authenticity enough to impact character evaluations; and it is possible that 
outsourcing to a personalized AI simply is not enough to overcome negative 
perceptions. While previous work has found an effect of personalized AI models 
on perceived credit [@Earp2024], and the majority of participants in our study 
stated that the personalized AI was more authentic than a standard model like 
ChatGPT, perceptions of *meaningful* authenticity in our study may have remained
low even with the personalized AI model. An AI could be perfectly trained on all
apologies that a person has ever written, but one might still think that a 
specific apology it then generates in a new instance is not an *authentic* 
apology. Therefore, even if people were described as outsourcing to an AI that 
was trained on their own writing and therefore personalized, participants still
may not have seen the specific output as being meaningfully authentic in the way
that matters for character judgments.

@@

Psychologists have long been interested in the apparent paradox of effort: it is 
both costly and valued [see @Inzlicht2018]. In more recent years, scholars have 
turned to consider not only why people might value effort itself, but how (not) 
expending effort shapes how we are viewed by others. @Celniker2023 argue that 
effort is moralized because people seem to infer positive moral character and 
cooperative intent from effort. But yet it has been unclear how outsourcing 
itself, and not reduced effort per se, may shape judgments, especially when
people outsource to artificial agents. Moreover, previous work looking at effort 
has tended to focus on a small number of examples that are not the most socially 
impactful: looking at how people judge the quality of products like paintings, 
poems, or suits of armor based on how much effort is perceived [@Kruger2004] or 
looking at how people judge others who expend more or less effort in employment 
and physical exercise [@Celniker2023] or, in the context of AI, those who use AI 
for work [@Reif2025]. But it has remained unclear how we might view others who 
outsource to AI; how these effects might vary based on how socio-relational the 
task is; how different ways of outsourcing and expending less effort influence 
perceptions; and how outsourcing has different effects on different kinds of 
social perceptions. And perhaps most importantly, through asking these 
questions, we gain insight into the most important question: why exactly does 
effort have the effects that it does? Is it that expending less effort directly 
leads to negative social perceptions, as some have argued in psychology 
[@Celniker2023] or, more in line with philosophical perspectives 
[e.g., @Aristotle2009] might expending effort also lead to second-order 
judgments about how much someone values a task, which then independently drives 
negative perceptions? Across our studies, we shed new light on all of these 
questions, providing a greater understanding of how, when, and why doing
something oneself the hard way shapes character judgments.

>>>>

---
title: "Negative Perceptions of Outsourcing to Artificial Intelligence"
shorttitle: "Outsourcing to AI"
author:
  - name: Scott Claessens
    corresponding: false
    orcid: 0000-0002-3562-6981
    email: scott.claessens@gmail.com
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - visualization
      - writing
      - editing
  - name: Pierce Veitch
    corresponding: false
    orcid: 0009-0005-3364-7470
    email: pv201@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - formal analysis
      - methodology
      - editing
  - name: Jim A.C. Everett
    corresponding: true
    orcid: 0000-0003-2801-5426
    email: j.a.c.everett@kent.ac.uk
    affiliations:
      - name: University of Kent
        department: School of Psychology
        address: Keynes College
        city: Canterbury
        country: UK
        postal-code: CT2 7NP
    role:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - editing
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: This pre-print is currently not yet peer-reviewed, and may differ from the final version.
    data-sharing: Word count (excluding abstract, figure/table legends, and references) = 10,408 words.
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "As artificial intelligence (AI) tools become increasingly integrated into daily life, people are beginning to outsource not only professional tasks but also socio-relational ones. Large language models like ChatGPT can generate wedding vows, speeches, and personal messages, raising questions about how individuals who use AI for such tasks are perceived by others. In this paper, we conduct six pre-registered studies with British participants (N = 3,935) to understand how people view those who expend less effort by outsourcing tasks to AI in different ways. We highlight a tradeoff between efficiency and inferred moral character, authenticity, and value: outsourcing makes us think more negatively about not only the person and their motivations, but also the outsourced work itself. Moreover, we show that this effect is not uniform but especially pronounced for more socio-relational tasks; that reduced effort does not lead to uniformly negative character perceptions across different tasks; and the reason that effort is important is because it signals second-order judgments that people are being less authentic and care less about the task. Our research highlights the way that AI use shapes our perceptions of people, raising key philosophical questions about efficiency, authenticity, and social ties in a world filled with AI-mediated interactions."
keywords: [artificial intelligence, person perception, outsourcing, effort, trust]
bibliography: bibliography.bib
engine: knitr
format:
  apaquarto-docx:
    fig-dpi: 300
    fig-align: center
execute:
  echo: false
  warning: false
  error: false
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
      caption-location: bottom
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
      caption-location: top
floatsintext: true
---

@@

The widespread release of generative AI language models has transformed daily 
life, offering the potential to perform a variety of tasks more efficiently and,
in some cases, with greater effectiveness than by doing them oneself. But as AI
becomes more widely available, people are not only using it to assist them with
things like preparing dinner recipes, writing data analysis code, and planning
daily schedules. Increasingly, AI might be used beyond routine or technical
domains to instead assist in tasks that are more socio-relational in nature,
like writing wedding vows, apology notes, and love letters. Anecdotal evidence
suggests that not only is AI-outsourcing of this kind already happening, but
that it potentially has serious effects on how we judge others. In a recent
Reddit post, a disgruntled newlywed tells the story of her husband using ChatGPT
to write his wedding vows, expressing her discomfort with outsourcing something
to AI that, to her, is deeply meaningful and a reflection of their love for one 
another [@miramar0]. Other reports describe people's negative reactions when
they learn that their romantic partner used ChatGPT to write them an apology 
note [@Tait2024] or even to break up with them [@Anderson2025]. Outsourcing 
tasks -- especially socio-relational ones -- to AI tools may be efficient, but 
this efficiency could trade off against inferred moral character, authenticity, 
and value.

In a world where the landscape of expending effort has changed through large
language models (LLMs), we need to know how people who outsource tasks to AI 
are perceived -- and why. We know that people are increasingly using LLMs for a 
wide variety of tasks [@UKGov2024]. Due to their ubiquity and their labelled 
intention as a tool for efficiency, perhaps outsourcing to LLMs might not not 
have consequences for perceived character? We are sceptical. We know that people
dislike it when others "free ride" or reduce effort while benefiting from 
collective resources [e.g. @Cubitt2011; @Kerr1983] and that people's outputs are
perceived as more valuable the more effort was ostensibly put into them 
[@Kruger2004]. And while there is much less attention paid in social psychology
to how expending effort shapes perceptions of character, the evidence there is 
suggests that exertion of effort is deemed morally admirable and is rewarded, 
even in situations where effort does not directly generate additional product, 
quality, or economic value, suggesting that effort itself is moralized 
[@Celniker2023]. While outsourcing and reduced effort are not synonymous, it 
seems the core social psychological processes are likely to be similar: when 
someone outsources a task instead of doing it themselves, they are expending 
less effort to achieve their goal, and people value effort. Indeed, the scant 
recent work that does exist suggests that describing someone as using AI for a 
relational task led to the perception they expended less effort and were less 
satisfied with their relationship [@Liu2024] and that people receive more 
negative social evaluations if they use AI to complete assignments at school and
at work [@Reif2025; @Roth2025]. We do not know, however, how outsourcing -- and 
therefore, presumably, expending less effort -- might be different for a human 
or an AI, and how these processes will occur in the myriad ways in which people 
can and do use AI to help them complete tasks.

Second, we need to know whether the *type* of task that people are outsourcing 
matters. One might expect outsourcing to be perceived negatively regardless of 
the type of task being outsourced -- if effort is generally moralized, then the 
domain in which it is expended (or not) should have little impact. However, 
there are reasons to expect differences between social tasks like writing vows 
and non-social tasks like writing computer code. We know that different norms,
standards, and expectations can be applied to social and non-social tasks and 
exchanges [e.g. @Fiske1992; @Heider1958; @Malle2022] and that what people want 
and expect from different kinds of relationships (and, therefore, associated 
tasks) varies [@Clark2012]. Indeed, from a philosophical perspective, it often 
matters not only *whether* something is done, but *how* it is done 
[@Aristotle2009; @Hursthouse2023; @Stohr2006]. An apology is not just about 
hearing someone say "I am sorry", but seeing genuine regret; a love letter is 
not just about hearing someone say "I love you", but seeing depth of emotion; 
and a bereavement letter is not just about hearing someone say "I am sorry for 
your loss", but seeing an understanding for the powerful human experience of 
loss. There is, perhaps especially for social tasks, value not only in the 
outcome of doing something, but the *process* too [@Goodman2010]. Previous work
on the role of effort, however, has seemed to suggest that effort should be 
uniformly important regardless of domain and has tended to focus on a limited 
number of tasks, looking at how people judge the quality of products like 
paintings, poems, or suits of armour based on how much effort is perceived 
[@Kruger2004] or how people judge others who expend more or less effort in 
employment and physical effort like marathon running [@Celniker2023]. Yet the 
number of tasks that people can outsource – and therefore, presumably, expend 
less effort on – is vast, and these tasks can vary in different ways that seem 
likely to have very different consequences for person perception. Different 
kinds of tasks that people can outsource can be more or less socially relevant, 
more or less consequential, and more or less likely to impact others: whether 
someone uses AI to spend less time planning a friend's birthday party or writing
an apology letter seems intuitively to be a different kind of thing than using
AI to spend less time writing a daily schedule or computer code to present a 
figure in an academic paper. To understand any potential negative effects of 
outsourcing -- to AI or to humans -- we must therefore look at a broad range of 
non-social and social tasks to understand not only potential generalizability 
but also *differences*.

Third, we need to know how different *ways* of outsourcing to AI influence 
negative perceptions. There are many different ways that people can use LLMs, 
and someone who "fully" outsources a task to AI by simply giving it a prompt and
copying the output word-for-word might be perceived very differently to someone
who gives the AI a prompt, revises the work accordingly, and finishes it 
themselves -- using AI as a *collaborative tool*, rather than as a replacement.
Similarly, someone could deceive others about their use of AI or be perfectly 
honest about it, and someone could use AI because they are in a rush or because
the task is important to them and they want to get it right. In all cases people
are expending less effort, but the way they do this is different and it remains
unclear how much this influences perceptions, and whether using AI in this way
is sufficient to eliminate negative perceptions: if someone does not do the task
fully themselves but uses an AI tool in the "best" way (by using it as a 
collaborative tool, being open about their use of AI, and having a good reason 
for using AI), would they still suffer negative social consequences from doing 
so?

Fourth, we need to understand *why* outsourcing to AI, and therefore expending 
less effort, might have these effects. Why is it that delegating to an AI might
have negative consequences for person perception – especially if people are, in
one sense, only using the AI tool in the precise way it has been designed and 
marketed? Previous work has focused on how expending less effort directly leads
to negative perceptions of others [@Celniker2023]. But this raises the question
of *why* effort is seen as important and what exactly it is signalling to
others, beyond one's general cooperative intent. It is possible that outsourcing
leads to negative perceptions because the lack of effort spent on the task
signals something more fundamental about how authentic one is and how much one
cares about the task: when someone chooses to outsource a love letter to an AI,
they might be seen as valuing that love letter and what it represents less. It
could be this second-step order of perceptions that is the key driver of
negative perceptions, and this, in turn, could especially be so for the kinds of
tasks that are more socially relevant.

@@

In this paper, we explore how people perceive others who outsource different 
kinds of tasks, in different ways, for different reasons, to AI. Across six 
pre-registered experiments with British participants, we seek not only to 
understand how reduced effort through AI-outsourcing might shape perceptions of 
others, but also advance previous theoretical work on the moralization of effort
[@Celniker2023]. Moving beyond a focus on how effort itself shapes character 
judgments uniformly across different domains, we seek to provide deeper 
theoretical nuance into why effort is important: that effort might be 
particularly important for socio-relational tasks, not only because effort 
itself is intrinsically moralized, but because expending effort gives rise to 
second-order inferences that independently influence character judgments about
authenticity and care.

@@

Interestingly, we found no effect of authenticity as proxied by the use of a 
personalized AI that is trained on one's own prior writings compared to a 
standard AI like ChatGPT. There are different explanations of this: it is 
possible that authenticity is not an important mechanism underlying the effect 
of outsourcing on negative character evaluations, but it could also suggest that
even using an personalized LLM still does not make someone truly authentic, in
the same way that we found even people who are honest and used AI as a tool are
still judged negatively for social tasks particularly. That is, while the 
majority of participants in our study stated that the personalized AI was more 
authentic than a standard model like ChatGPT, using a personalized LLM is still
not meaningfully authentic because of its nature: that this is a feature of the
phenomenon, not a bug. An AI could be perfectly trained on all apologies that a 
person has ever written, but one might still think that a specific apology it 
then generates in a new instance is not an *authentic* apology because it does 
not come from the kind of processes deemed important in an apology: a personal 
recollection of the wrong, a commitment to change. An apology written by an AI, 
even in someone's own style, is still arguably not an actual apology.

@@

There is an apparent paradox of effort: it is both costly and valued [see 
@Inzlicht2018]. While seemingly critical to our social lives, though, there has 
been relatively little attention paid to how (not) expending effort shapes how 
we are viewed by others. @Celniker2023 argue that effort is moralized because 
people seem to infer positive moral character and cooperative intent from 
effort. Yet this work has looked only at a few types of cases where effort is 
expended or people seek outside help, where these are not necessarily the most 
socially impactful: looking at how people judge the quality of products like 
paintings, poems, or suits of armor based on how much effort is perceived 
[@Kruger2004] or looking at how people judge others who expend more or less 
effort in employment and physical exercise [@Celniker2023] or, in the context of
AI, those who use AI for work [@Reif2025]. It has remained unclear how we might
view others who outsource to AI specifically; how these effects might vary based
on how socio-relational the task is; how different ways of outsourcing and 
expending less effort influence perceptions; and how outsourcing has different
effects on different kinds of social perceptions. And perhaps most importantly,
through asking these questions, we gain insight into the most important
question: why exactly does effort have the effects that it does? Why do people
care not only about the outcome achieved but the process by which it is done,
and why might outsourcing to AI (or humans) trade off efficiency for inferred
authenticity, care, and moral character? Is it that expending less effort 
directly leads to negative social perceptions, as some have argued in psychology 
[@Celniker2023] or, more in line with philosophical perspectives 
[e.g., @Aristotle2009] might expending effort also lead to second-order 
judgments about how much someone values a task, which then independently drives 
negative perceptions? Across our studies, we shed new light on all of these 
questions, providing a greater understanding of how, when, and why doing
something oneself the hard way shapes character judgments.
